---
title: "Science Communication with R"
author: "Antoine Bichat"
date: "State Of the R - AgroPariTech<br>February 23, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(tidyverse)
```

class: inverse, center, middle

# Code presentation


---
# Syntax


* Use `<-` for assignation and `=` for arguments

* Leave a space

  * before `(` (except when it’s part of a function call)

  * after `)` or `,` 

  * around `=`, `<-`, `+`, `*`, `==`, ...
  
```{r, fig.height=3}
x <- rnorm(n = 100000)^2 + rnorm(n = 100000)^2
hist(x[x < 10], col = "grey", xlab = "", ylab = "Count", 
     main = "Chi-squared empirical distribution")
```


---
# Functions names
 
* When possible, use verbs for function names: `add_row()` or instead of `row_adder()`

--

* A function should do one thing well (it should be called either because it has side-effects or because it returns a value; not both)

--

* Don't hesitate to split your long function in smaller ones.

--

* Only use `return()` for early returns.

--

* Comment your functions (use the `Ctrl + Maj + C shortcut`)

---
# Pipes

* For long pipes, use one verb per line

* Don't ommit the brackets

```{r eval=FALSE}
# Good
summary_SL <-
  iris %>% 
  drop_na() %>% 
  group_by(Species) %>% 
  summarise(Mean_SL = mean(Sepal.Length))

# Bad
summary_SL <- iris %>% drop_na %>% group_by(Species) %>% 
  summarise(Mean_SL=mean(Sepal.Length))

iris %>% 
  drop_na %>% 
  group_by(Species) %>% 
  summarise(Mean_SL=mean(Sepal.Length)) -> summary_SL
```


---
# Useful help

* The Ctrl + Maj + A shortcut rearranges your code

--

* {styler} package


```{r slyler}
ugly_code <- "a<-function( x){1+1}           "
styler::style_text(ugly_code)
```



---
class: middle, center, inverse

# References


---
## Style

* The tidyverse style guide http://style.tidyverse.org

## Rmarkdown

---
class: center, middle, inverse

# Thanks for your attention!

.footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).]
