---
title: "Visualization with R (and ggplot2)"
author: "Antoine Bichat / Erwan Le Pennec"
date: "Spring 2019"
output:
   html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r Knitr_Global_Options, include=FALSE}
library("knitr")
opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE, autodep = TRUE, tidy = FALSE)
library("rgl")
knit_hooks$set(webgl = hook_webgl)
```

```{r Tidyverse}
library(tidyverse)
```


# One variable graph

## Bar

```{r Bar}
ggplot(data = diamonds, aes(x = cut, fill = cut)) +
  geom_bar()
```

## Pie plot

```{r Pie}
ggplot(data = diamonds, aes(x = factor(1), fill = cut)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y") +
  xlab(NULL) + ylab(NULL) +
  theme(
    axis.ticks = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r Pie_Ggforce}
ggplot(data = diamonds %>% count(cut)) +
  ggforce::geom_arc_bar(
    aes(
      x0 = 0, y0 = 0,
      r0 = 0, r = 1,
      amount = n, fill = cut
    ),
    stat = "pie"
  ) +
  coord_equal() + theme_void()
```

```{r Pie_Ggforce_function}
library(rlang)
pie_stats <- function(df, x0, y0, r0, r1, amount, explode, label_perc) {
  x0 <- enquo(x0)
  y0 <- enquo(y0)
  r0 <- enquo(r0)
  r1 <- enquo(r1)
  amount <- enquo(amount)
  explode <- enquo(explode)

  df %>%
    mutate(
      `x0` = !! x0,
      `y0` = !! y0,
      `r0` = !! r0,
      `r1` = !! r1,
      `explode` = !! explode
    ) %>%
    group_by(x0, y0) %>%
    mutate(end = cumsum(!! amount) / sum(!! amount) * 2 * pi) %>%
    mutate(start = lag(end, default = 0)) %>%
    ungroup() %>%
    mutate(
      x_lab = (!! x0) +
        ((!! r0) + label_perc * ((!! r1) - (!! r0)) + (!! explode)) *
          sin((end + start) / 2),
      y_lab = (!! y0) +
        ((!! r0) + label_perc * ((!! r1) - (!! r0)) + (!! explode)) *
          cos((end + start) / 2)
    )
}
```

```{r Pie_Ggforce_Legend}
ggplot(data = diamonds %>%
  count(cut) %>%
  pie_stats(0, 0, 0, 1, n, .2 * (cut == "Fair"), .8)) +
  ggforce::geom_arc_bar(
    aes(
      x0 = x0, y0 = y0, r0 = r0, r = r1,
      start = start,
      end = end,
      explode = explode,
      fill = cut
    )
  ) +
  geom_text(
    aes(
      x = x_lab, y = y_lab,
      label = scales::percent(n / sum(n))
    ),
    size = 3
  ) +
  coord_equal() + theme_void()
```

```{r Col_vs_Pie}
set <- tibble(cat = c("A", "B", "C", "D", "E"))
sets <- purrr::map_df(
  1:5,
  function(i) {
    set %>% mutate(
      val = rpois(5, 50),
      rep = i
    )
  }
)

p_col <- ggplot(data = sets) +
  geom_col(
    aes(x = cat, y = val, fill = cat),
    show.legend = FALSE
  ) +
  facet_wrap(~ rep, ncol = 5)

p_pie <- ggplot(data = sets) +
  ggforce::geom_arc_bar(
    aes(
      x0 = 0, y0 = 0,
      r0 = 0, r = 1,
      amount = val, fill = cat
    ),
    stat = "pie", show.legend = FALSE
  ) + facet_wrap(~ rep, ncol = 5) + theme_void()

egg::ggarrange(p_pie, p_col, ncol = 1, 
               heights = c(.4, 1))
```

## Cleveland Dot

```{r Cleveland_Dot}
ggplot(data = diamonds, aes(x = cut)) +
  geom_point(stat = "count", size = 5) + coord_flip() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(
      linetype = "dotted",
      color = "darkgray"
    )
  )
```

```{r Lollipop}
ggplot(data = diamonds, aes(x = cut, color = cut,
                            fill = cut)) +
  geom_point(stat = "count", size = 5) +
  geom_bar(width = .01) + coord_flip() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(
      linetype = "dotted",
      color = "darkgray"
    )
  )
```

```{r Cleveland_Ordered}
ggplot(
  data = rownames_to_column(mtcars),
  aes(x = fct_reorder(rowname, mpg),
      y = mpg, color = cyl)) +
  geom_point(size = 3) + coord_flip() + xlab("") +
  scale_color_continuous(breaks = c(4, 6, 8)) +
  guides(color = guide_legend()) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(
      linetype = "dotted",
      color = "darkgray")
  )
```

## Histogram and density

```{r Histogram}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram()
```

```{r Histogram_Oversmoothing}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram(binwidth = 15)
```

```{r Histogram_Undersmoothing}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram(binwidth = 100)
```

```{r Density}
ggplot(data = diamonds, aes(x = price)) + geom_density()
```

```{r Density_Oversmooting}
ggplot(data = diamonds, aes(x = price)) +
  geom_density(adjust = 4)
```


```{r Density_Undersmooting}
ggplot(data = diamonds, aes(x = price)) +
  geom_density(adjust = .01)
```

## Boxplot and violin plot

```{r Boxplot}
ggplot(data = diamonds, aes(x = factor(1), y = price)) +
  geom_boxplot()
```

```{r Violin}
ggplot(data = diamonds, aes(x = factor(1), y = price)) +
  geom_violin()
```

## Dot plot

```{r Dotplot_Data}
small_diamonds <- sample_frac(diamonds, .01)
```

```{r Dotplot}
ggplot(data = small_diamonds,
       aes(x = factor(1), y = price)) +
  geom_dotplot(
    binaxis = "y", stackdir = "center",
    stackratio = 1.25, dotsize = .25
  )
```

```{r Doplot_Median}
ggplot(data = small_diamonds,
       aes(x = factor(1), y = price)) +
  geom_dotplot(
    binaxis = "y", stackdir = "center",
    stackratio = 1.25, dotsize = .25
  ) +
  stat_summary(
    fun.y = median, geom = "point",
    size = 5, color = "red"
  )
```

```{r Quasirandom}
ggplot(small_diamonds, 
       aes(x = factor(1), y = price)) +
  ggbeeswarm::geom_quasirandom()
```

## Stem

```{r Stem}
stem(small_diamonds[["price"]], scale = 3)
```

## Grouping

```{r Grouping_Color}
ggplot(data = diamonds, aes(x = price, color = cut)) +
  geom_density()
```

```{r Grouping_Fill_Stack}
ggplot(data = diamonds, aes(x = price, fill = cut)) +
  geom_density(position = "stack")
```

```{r Grouping_Facet}
ggplot(data = diamonds, aes(
  x = price, color = cut,
  fill = cut
)) +
  geom_density() + facet_wrap(~ cut)
```

```{r Grouping_Violin}
ggplot(data = diamonds, aes(x = cut, y = price)) +
  geom_violin()
```

```{r Grouping_Ridges}
ggplot(data = diamonds, aes(y = cut, x = price)) +
  ggridges::geom_density_ridges()
```

```{r Grouping_Dotplot}
ggplot(data = small_diamonds, aes(x = cut, y = price)) +
  geom_dotplot(
    binaxis = "y", stackdir = "center",
    stackratio = 1.25, dotsize = .25
  ) +
  stat_summary(
    fun.y = median, geom = "point",
    size = 5, color = "red"
  )
```

```{r Grouping_Dotplot_Color}
ggplot(
  data = small_diamonds,
  aes(x = factor(1), y = price, color = cut)
) +
    geom_dotplot(
      aes(fill = cut),
      binaxis = "y", stackdir = "center",
      stackratio = 1.25, dotsize = .25,
      binpositions = "all", stackgroups = TRUE
    ) +
  stat_summary(
    fun.y = median, geom = "point",
    size = 5, color = "red",
    show.legend = FALSE
  )
```

# Relationship graph

## Scatter Plots

```{r Scatter_Plot}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()
```

```{r Scatter_Plot_Alpha}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = .25)
```

```{r Scatter_Plot_Color}
ggplot(data = diamonds, aes(
  x = carat, y = price,
  color = cut
)) +
  geom_point()
```

```{r Scatter_Plot_Color_Alpha}
ggplot(data = diamonds, aes(
  x = carat, y = price,
  color = cut
)) +
  geom_point(alpha = .25)
```

```{r Scatter_Plot_Color_Alpha_Facet}
ggplot(data = diamonds, aes(
  x = carat, y = price,
  color = cut
)) +
  geom_point(alpha = .25) + facet_wrap(~ cut)
```

## Smoothing

```{r Smoothing}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()
```

```{r Smoothing_Group}
ggplot(data = diamonds, aes(
  x = carat, y = price,
  color = cut
)) +
  geom_point(alpha = .25) + geom_smooth()
```

## Symbols

```{r Symbols}
ggplot(data = mtcars, aes(
  x = hp, y = mpg,
  size = gear, color = cyl,
  shape = factor(am)
)) +
  geom_point() +
  scale_size_continuous(
    breaks = c(3, 4, 5),
    limits = c(0, 5)
  ) +
  scale_color_continuous(breaks = c(4, 6, 8)) +
  guides(color = guide_legend())
```

## Scatter Plot Matrix

```{r Scatter_Plot_Matrix}
library(GGally)
ggpairs(mtcars)
```

## Parallel Coordinates and Radar Plots

```{r Parallel_Coordinates_Data}
# rescale all variables to lie between 0 and 1
scaled <- as.data.frame(lapply(mtcars, ggplot2:::rescale01)) %>%
  mutate(model = row.names(mtcars))
# add model names as a variable
mtcarsm <- gather(scaled, variable, value, -model)
```

```{r Parallel_Coordinates}
ggplot(mtcarsm, aes(x = variable, y = value)) +
  geom_line(aes(group = model, color = model), size = 2) +
  theme(
    strip.text.x = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.8))
  ) +
  guides(color = guide_legend(ncol = 2))
```

```{r Parallel_Coordinates_Facet}
ggplot(mtcarsm, aes(x = variable, y = value)) +
  geom_line(aes(group = model, color = model), size = 2) +
  theme(
    strip.text.x = element_text(size = rel(0.8)),
    axis.text.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  guides(color = "none") + facet_wrap(~ model, nrow = 4)
```

```{r Radar_Plots_Function}
coord_radar <- function(theta = "x", start = 0, direction = 1) {
  theta <- match.arg(theta, c("x", "y"))
  r <- if (theta == "x") {
    "y"
  } else {
    "x"
  }
  ggproto(
    "CordPolar", CoordPolar, theta = theta, r = r, start = start,
    direction = sign(direction),
    is_linear = function(coord) TRUE
  )
}
```

```{r, Radar_Plot_Facet}
ggplot(mtcarsm %>% dplyr::arrange(variable), aes(x = variable, y = value)) +
  geom_polygon(aes(group = model, color = model), fill = NA, size = 1, show.legend = FALSE) +
  coord_radar() +
  facet_wrap(~ model, nrow = 4) +
  guides(color = guide_legend(ncol = 2)) +
  theme(
    strip.text.x = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.8)),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```


```{r, Radar_Plot}
ggplot(mtcarsm %>% dplyr::arrange(variable), aes(x = variable, y = value)) +
  geom_polygon(aes(group = model, color = model), fill = NA, size = 2, show.legend = FALSE) +
  geom_path(aes(group = model, color = model), size = 2) +
  coord_radar() +
  theme(
    strip.text.x = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.8))
  ) +
  guides(color = guide_legend(ncol = 2))
```

# Time


## Time Series

```{r Time_Series_Data}
tckrs <- c("SPY", "QQQ", "GDX", "DBO", "VWO")

stocks_tbl <- tidyquant::tq_get(tckrs, from = "2007-01-01")

stocks_tbl <- stocks_tbl %>% group_by(symbol) %>%
  mutate(close_index = close/close[1])
```

```{r Time_Series}
ggplot(stocks_tbl, aes(x = date, y = close_index, color = symbol)) +
  geom_line() +
  scale_y_continuous(limits = c(0, NA))
```

```{r Time_Series_Facet}
ggplot(stocks_tbl, aes(x = date, y = close_index,
                       color = symbol)) +
  geom_line() +
  scale_y_continuous(limits = c(0, NA)) + 
  facet_wrap(~ symbol)
```

## Time Line

```{r Time_Line_Data}
library(timeline)
data(ww2)
```

```{r Time_Line}
timeline(ww2, ww2.events, event.spots = 2, event.label = "", event.above = FALSE)
```

# Map

## Choroplet

```{r Choroplet_Data}
st <- map_data("state")
data(votes.repub)
colnames(votes.repub) <- paste("Y", colnames(votes.repub), sep = "")
votes.repub <- mutate(rownames_to_column(as.data.frame(votes.repub), "region"), region = tolower(region))
st <- left_join(st, votes.repub, by = "region")

st_gather <- gather(
  st, "year", "value",
  -long, -lat, -group, -order, -region, -subregion
)
```

```{r Choroplet}
ggplot(st, aes(long, lat, group = group, fill = Y1960)) +
  geom_polygon() +
  scale_fill_gradient(limits = c(20, 80), low = "blue", high = "red") +
  coord_map()
```

```{r Choroplet_Year}
ggplot(st_gather, aes(long, lat, group = group, fill = value)) +
  geom_polygon() +
  scale_fill_gradient(limits = c(20, 80), low = "blue", high = "red") +
  facet_wrap(~ year) +
  coord_map()
```

```{r Choroplet_Clean}
pacman::p_unload(maps)
```

# Graph

## Alluvial

```{r Alluvial}
library(ggalluvial)
ggplot(
  data = as.data.frame(Titanic),
  aes(
    weight = Freq, axis1 = Class,
    axis2 = Sex, axis3 = Age,
    axis4 = Survived
  )
) +
  scale_x_continuous(breaks = 1:4,
                     labels = c("Class", "Sex",
                                "Age", "Survived")) +
  geom_alluvium(aes(fill = Survived), reverse = FALSE) +
  geom_stratum(reverse = FALSE) +
  geom_text(stat = "stratum", label.strata = TRUE, reverse = FALSE)
```

## Tree

```{r Tree_Data}
library(rpart)
data(airquality)
airq <- subset(airquality, !is.na(Ozone))
airct <- rpart(Ozone ~ ., data = airq,
               control = rpart.control(minsplit = 10))
```

```{r Tree_Rpartplot}
library(rpart.plot)
rpart.plot(airct)
prp(airct, type = 2, extra = 1, nn = TRUE)
prp(airct, type = 2, extra = 1, nn = TRUE, fallen.leaves = TRUE)
```

```{r Tree_Ggdendro}
library(ggdendro)
airct.data <- dendro_data(airct)
ggplot() +
  geom_segment(
    data = airct.data$segments,
    aes(x = x, y = y, xend = xend, yend = yend)
  ) +
  geom_label(
    data = airct.data$labels,
    aes(x = x, y = y, label = label), 
    size = 3, vjust = .5
  ) +
  geom_text(
    data = airct.data$leaf_labels,
    aes(x = x, y = y, label = label), size = 3, vjust = 1.5
  ) +
  theme_dendro()
```

## Tree Graph

```{r Tree_Graph_Data}
data(business, package = "treemap")
```

```{r Tree_GGraph_Data}
library(ggraph)
library(tidygraph)
edge0 <- business %>% group_by(from = "0", to = NACE1) %>% summarize()
node0 <- business %>%
  group_by(name = "0") %>%
  summarize(employees = sum(employees, na.rm = TRUE))
edge1 <- business %>% group_by(from = NACE1, to = NACE2) %>% summarize()
node1 <- business %>%
  group_by(name = NACE1) %>%
  summarize(employees = sum(employees, na.rm = TRUE))
edge2 <- business %>%
  group_by(from = NACE2, to = NACE3) %>%
  summarize()
node2 <- business %>%
  group_by(name = NACE2) %>%
  summarize(employees = sum(employees, na.rm = TRUE))
edge3 <- business %>%
  group_by(from = NACE3, to = NACE4) %>%
  summarize()
node3 <- business %>%
  group_by(name = NACE3) %>%
  summarize(employees = sum(employees, na.rm = TRUE))
node4 <- business %>%
  group_by(name = NACE4) %>%
  summarize(employees = sum(employees, na.rm = TRUE))
business_edge <- bind_rows(edge0, edge1, edge2, edge3)
business_node <- bind_rows(node0, node1, node2, node3, node4) %>%
  group_by(name) %>%
  summarize(employees = 1 + sum(employees, na.rm = TRUE)) %>%
  mutate(name2 = str_split(name, "-", 2) %>% map_chr(tail, 1))
business_igraph <- igraph::graph_from_data_frame(
  business_edge,
  vertices = business_node
) %>%
  as_tbl_graph() %>%
  activate("nodes") %>%
  mutate(name3 = if_else(node_is_sink() & (employees > 1000),
    name2, ""
  ))

```

```{r Tree_GGraph_Treegraph}
ggraph(business_igraph, layout = "dendrogram",
       circular = TRUE) + 
  geom_edge_fan(alpha = .5) + 
  geom_node_point(aes(color = name),
                  show.legend = FALSE) +
  coord_equal() +
  theme_graph()
```

```{r Tree_GGraph_Treemap}
ggraph(business_igraph, layout = "treemap",
       weight = "employees") +
  geom_node_tile(aes(fill = name), show.legend = FALSE) +
  geom_node_text(
    aes(label = str_trunc(str_trim(name3), 10)),
    size = 2,
    show.legend = FALSE) +
  theme_graph()
```

## General Graph

```{r Graph_Data}
library(tidygraph)
mis_file <- "lesmiserables.txt"
mis_graph <- igraph::read_graph(mis_file,
                                format = "gml") %>%
  as_tbl_graph()

mis_graph <- mis_graph %>% activate("nodes") %>%
  mutate(group = as_factor(as.character(group)),
         degree = centrality_degree()) %>%
  arrange(group, desc(degree))
```

```{r GGraph}
library(ggraph)
ggraph(mis_graph) + geom_edge_fan(
  aes(edge_width = value),
  alpha = .5,
  show.legend = FALSE
) +
  geom_node_point(
    aes(color = as.factor(group)),
    size = 5,
    show.legend = FALSE
  ) +
  geom_node_text(aes(label = label)) +
  scale_x_continuous(expand = c(0.1,0.1)) +
  coord_equal() +
  theme_graph()

ggraph(mis_graph, layout = "fr") +
  geom_edge_fan(
    aes(edge_width = value),
    alpha = .5,
    show.legend = FALSE
  ) +
  geom_node_point(
    aes(color = as.factor(group)),
    size = 5,
    show.legend = FALSE
  ) +
  geom_node_text(aes(label = label)) +
  scale_x_continuous(expand = c(0.1,0.1)) +
  coord_equal() +
  theme_graph()

ggraph(mis_graph, layout = "circle") +
  geom_edge_fan(
    aes(edge_width = value),
    alpha = .5,
    show.legend = FALSE
  ) +
  geom_node_point(
    aes(color = as.factor(group)),
    size = 5,
    show.legend = FALSE
  ) +
  geom_node_text(aes(label = label)) +
  scale_x_continuous(expand = c(0.1,0.1)) +
  coord_equal() +
  theme_graph()
```

## Arc diagram

```{r GGraph_Arcdiagram}
ggraph(mis_graph, layout = "linear") +
  geom_edge_arc(
    aes(edge_width = value),
    alpha = .25,
    show.legend = FALSE
  ) +
  geom_node_point(
    aes(
      color = as.factor(group),
      size = degree
    ),
    show.legend = FALSE
  ) +
  geom_node_text(
    aes(label = label), angle = 90,
    hjust = 1
  ) +
  scale_y_continuous(limits = c(-10, NA)) +
  theme_graph()
```

##  Adjacency Matrix

```{r Adjacency_Matrix_Data}
edges <- mis_graph %>% activate("edges") %>% as_tibble()
edges_sym <- edges %>% rename(tmp = from, from = to) %>%
  rename(to = tmp)
edges <- bind_rows(edges, edges_sym)
edges <- edges %>%
  left_join(mis_graph %>% activate("nodes") %>% as_tibble() %>%
              select(id, from_label = label),
            by = c("from" = "id")) %>%
  left_join(mis_graph %>% activate("nodes") %>% as_tibble() %>%
              select(id, to_label = label),
            by = c("to" = "id")) %>%
  mutate(from_label = fct_reorder(from_label, from),
         to_label = fct_reorder(to_label, to))
```

```{r Adjacency_Matrix}
ggplot(data = edges, aes(
  x = from_label,
  y = to_label,
  fill = value
)) +
  geom_raster() + xlab("") + ylab("") +
  viridis::scale_fill_viridis(trans = "sqrt") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Interactivity


## OpenGL

```{r 3D, webgl = TRUE}
library(rgl)

plot3d(small_diamonds[c(1, 2, 7)])
```


## Javascript

### Plotly 

```{r Plotly, cache = FALSE}
library(plotly)
gg <- ggplot(data = mtcars, aes(
  x = hp, y = mpg,
  size = gear, color = cyl,
  shape = factor(am)
)) +
  geom_point() +
  scale_size_continuous(
    breaks = c(3, 4, 5),
    limits = c(0, 5)
  ) +
  scale_color_continuous(breaks = c(4, 6, 8))
```

```{r Plotly2, cache = FALSE}
library(plotly)
ggplotly(gg)
```

### Bokeh

```{r Bokeh, cache = FALSE}
library(rbokeh)
p <- figure() %>% 
  ly_points(data = dplyr::mutate(mtcars, cyl = factor(cyl)),
            x = hp, y = mpg, color = cyl,
            glyph = factor(am), hover = mtcars)
p
```

```{r Bokeh_advanced, cache = FALSE}
p <- figure(data = stocks_tbl) %>%
  ly_lines(x = date, y = close_index, color = symbol)
p

ps <- purrr::map(tckrs,
                 function(tckr) {
                   figure(data = stocks_tbl %>%
                            dplyr::filter(symbol == tckr)) %>%
                     ly_lines(x = date, y = close, color = symbol)
                 })
grid_plot(ps, same_axes = c(TRUE,FALSE),
          ncol = 1)
```

### visNetwork

```{r visNetwork, cache = FALSE}
library(visNetwork)

visIgraph(mis_graph, idToLabel = FALSE,
          physics = TRUE) %>%
  visOptions(highlightNearest = TRUE) %>%
  visPhysics(solver = "barnesHut")
```

# Big Data

```{r BigData}
library(nycflights13)
glimpse(flights)

ggplot(data = flights,
       aes(x = distance, y = air_time)) +
  geom_point(alpha = .1)

ggplot(data = flights,
       aes(x = distance, y = air_time)) +
  geom_hex(bins = 60) +
  viridis::scale_fill_viridis()

ggplot(data = flights,
       aes(x = distance, y = air_time)) +
  stat_density_2d(aes(fill = ..level..),
                  contour = TRUE, geom = "polygon") +
  viridis::scale_fill_viridis()
```